\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-07-03 write_up/mixfishsim_ecolmod.tex
Text update

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%%%%%%%%%%%%%%%%%%%%%%%

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions
and using species non-selective fishing gear. \added[id=PD]{In doing so}
fisheries \deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011}.  \\

@@

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}. Each population \replaced[id=PD]{was}{is} initialised at
a single location, and subsequently move\replaced[id=PD]{ed}{s} according to a
probabilistic distrib\added[id=PD]{u}tion based on habitat suitability
(represented by the normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these element implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution (\cite{Bailey2018}). The decision to use an individual
based model for fishing vessels was taken because fishers as a group tend to
show heterogeneity and individual rather than group dynamics
[\textcolor{red}{REF!}]. Thus this was the fleet dynamics is the productive of
individual experiences rather than pre-ordained group dynamics.

@@

The survey simulation was set up with follow a fixed gridded station design
with 100 stations fished each year, starting on day 92 \added[id=PD]{and ending
	on day 112 (5 stations per day)} with same catchability parameters for
all populations ($Q_{p} = 1$). This was so as to approximate a real world
survey design with limited seasonal and spatial coverage.

@@

Our approach captures fine scale population and fishery dynamics and their
interaction in a way not usually possible with real data and thus not usually
considered in fisheries simulations. While other simulation frameworks seek to
model individual vessel dynamics based on inferred dynamics from VMS and
logbook records \citep{Bastardie2010}, or as a system to identify measures to
meet particular management goals \citep{Bailey2018}, our framework allows users
to explore the assumptions in modelling observations data and evaluate the
underlying dynamics of such approaches at a fine spatial and temporal scale.
This offers the advantage that larger scale fishery patterns are emergent
properties of the system rather and results can be compared to those obtained
under a statistical modelling framework. \\

@@

Survey design \\

commercial index standardization methods \\

>>>>

%% Harvard
\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%%%%%%%%%%%%%%%%%%%%%%%
%% correct equation line numbering

\let\oldequation\equation
\let\oldendequation\endequation
\renewenvironment{equation}
 {\linenomathNonumbers\oldequation}
 {\oldendequation\endlinenomath}

%%%%%%%%%%%%%%%%%%%%%%

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions
and using species non-selective fishing gear. \added[id=PD]{In doing so}
fisheries \deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011}\textcolor{red}{(ADD COSGROVE REFERENCE HERE)}.  \\

@@

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, which simulates a Gaussian Random Field process given a
user defined error model and correlation structure. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population
\replaced[id=PD]{was}{is} initialised at a single location, and subsequently
move\replaced[id=PD]{ed}{s} according to a probabilistic
distrib\added[id=PD]{u}tion based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these elements implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution (\cite{Bailey2018}). The decision to use an individual
based model for fishing vessels was taken because fishers as a group tend to
show heterogeneity and individual rather than group dynamics
[\textcolor{red}{REF!}]. Therefore in the simulations fleet dynamics are the
productive of individual experiences rather than pre-defined group dynamics.

@@

The survey simulation was set up with a fixed gridded station design with 100
stations fished each year, starting on day 92 \added[id=PD]{and ending on day
	112 (5 stations per day)} with same catchability parameters for all
populations ($Q_{p} = 1$). This approximates a real world survey design with
limited seasonal and spatial coverage.

@@

Our approach captures fine scale population and fishery dynamics and their
interaction in a way not usually possible with real data and thus not usually
considered in fisheries simulations. While other simulation frameworks seek to
model individual vessel dynamics based on inferred dynamics from VMS and
logbook records \citep{Bastardie2010}, or as a system to identify measures to
meet particular management goals \citep{Bailey2018}, our framework allows users
to explore the assumptions in modelling observational data and evaluate the
underlying dynamics of such approaches at a fine spatial and temporal scale.
This offers the advantage that larger scale fishery patterns are emergent
properties of the system rather and results can be compared to those obtained
under a statistical modelling framework. \\

COMPARE AGAINST OTHER MODELS: Osmose, Atlantis,... do they use rationale actors
and economic profit making ?

@@

Survey design - random, fixed, stratified - all possible to explore  \\

Assumptions in spatiotemporal models: GAMs, GLMMs, geostatistical models

Natural mortality estimation

Commercial index standardization methods - different spatial scales, individual
vessels, data aggregation -- all these can be explored. \\
